Meta-learning in movement prediction problem of aperiodic time-series data

### 2024-10-08 doc/paper/paper.tex

    \textbf{Abstract}. Predicting aperiodic time-series data (e.g. stock price, foreign exchange, Bitcoin price,...) is a difficult task for machine learning models because this type of data has high and not stationary variance; does not show clearly periodicity, making it difficult to extract features; depends not only on past values but also on external factors, which make them unstable and non-cyclical, such as economic and political situations. To overcome the above challenges, we use Meta-learning to train a combined \verb|LSTM| and \verb|CNN| network, thereby effectively extracting and synthesizing hidden features of the data over time. Experiments on foreign exchange data of 60 currency pairs over 24 years (2000-2024) show that the proposed method performs well and has 5\%-11\% higher accuracy compared to the \verb|NHITS| - the state-of-the-art model in 2023 on time-series data, in the task of predicting the trend (upward or downward) of the next trading day.

@@

For the variance variation challenge, ensemble models \citep{sadeghi2021combined, zafeiriou2020intraday, ali2020complete} are often used to mitigate the effects of variance variation. Ensemble models provide a holistic and multi-perspective view based on sub-models, thereby helping the overall model adapt to strong variance changes. However, under the rigid synthesis of traditional ensemble models, the results obtained on non-periodic datasets are very poor.

@@

To overcome the pattern variation challenge, most studies use features extracted from Long Short-Term Memory neural network (\verb|LSTM|) \cite{hochreiter1997long}, Artificial Neural Network (\verb|ANN|), and Convolution Neural Network (\verb|CNN|) \cite{lecun1989handwritten}. Specifically, in 2022, 20\% of all publications related to financial index prediction used \verb|LSTM|, 20\% used \verb|ANN|, and 6\% used \verb|CNN| \cite{ayitey2023forex}. This is completely understandable because \verb|LSTM| and \verb|CNN| have proven their ability to extract spatial and temporal features effectively. However, when faced with aperiodic data with only 2600 samples, after 40 training steps, \verb|LSTM| is still underfitting, while \verb|CNN| is overfitting after only 5 training steps.

@@

Regarding the external influences challenge, \cite{fama1970efficient} hypothesizes that different time-series datasets of the same domain at the same time reflect the impact of external factors. For example, studies \citep{overreactioncontrarian, mech1993portfolio} show the dependence between the financial ratios of a given company and the ratios of other companies. This further strengthens the hypothesis in study \cite{fama1970efficient}. Therefore, by exploiting data sources with the same domain, we can capture information reflecting external factors.

@@

We approach the above challenges through combination of feature combination and model synthesis. Specifically, by using Meta-learning (ML) \cite{finn2017model}, we efficiently synthesize the parameters of local models, which significantly reduces variance loss and increases the ability to collect information reflecting external factors from datasets in the same domain. To effectively learn laten patterns, we propose a method that combines features from \verb|CNN| and \verb|LSTM|. In addition, we assume that aperiodic time-series data also has hidden long-term dependencies. Traditional approaches use a fixed amount of past data (lookback window) to train the model. This poses a major obstacle to the learning process because long-term features will be forgotten over time. On the other hand, ML divides the dataset into many parts to learn and synthesize the learned parameters effectively, so it can handle this challenge well.

@@

Finally, we demonstrate the superiority of the proposed algorithm by solving the problem of predicting the trend (upward or downward) of foreign exchange rates and comparing the results with the existing state-of-the-art (SOTA) model (\verb|NHITS| \cite{challu2023nhits}) on two types of data: (1) - USD/JPY exchange rate data; (2) - Exchange rate data of 60 currency pairs, comprising 18 countries. These data sets are publicly available on the Internet and can be easily downloaded. The official implementation can be found at \url{https://github.com/baolongnguyenmac/multi_fx}.

@@

In summary, our main contributions include the following: First, we present a novel \textbf{feature combination technique} for aperiodic time-series data that extracts and combines features using both \verb|LSTM| and \verb|CNN| models. Second, we uncover hidden long-term dependencies, demonstrating experimentally that aperiodic time-series data not only rely on external factors but also exhibit dependencies with past period, which we call \textbf{hidden long-term dependencies}. Third, we introduce an \textbf{efficient model parameter aggregation approach} by employing ML algorithm in place of traditional ensemble models for combining parameters from multiple machine learning models. Finally, we validate our method through experiments on exchange rate datasets and show its effectiveness in comparison to the SOTA model, \verb|NHITS|.

@@

As aforementioned, \verb|LSTM| is a well-known neural network for handling prediction problems on time-series data. \verb|LSTM| is commonly used because it handles the problem of vanishing gradients well (easily encountered when using Recurrent neural networks) and can effectively exploit nonlinear relationships in data. Indeed, by maintaining the cell-state in each iteration, \verb|LSTM| can overcome the vanishing gradient problem, thereby preserving the ability to capture long-term dependencies \cite{cheng2018applied}. In addition, \verb|LSTM| performs feature extraction with nonlinear activation functions, which helps the model parameters capture the nonlinearity of the data \cite{he2016exploiting}. These factors make \verb|LSTM| to be the first choice to think of when solving problems on time-series data.

@@

Hybrid ensemble models have been widely used in time-series processing problems and have been experimentally proven to be more accurate than standard time-series models because they can synthesize the strengths of many sub-models \cite{ayitey2023forex}. However, current ensemble model synthesis forms are still very rigid because they can only synthesize based on the final results (e.g. voting mechanism of bagging models) and near-final results (e.g. stacking models). From the perspective of ensemble models, the equation \ref{eq:outer_opt} can be considered an effective method of synthesizing sub-models, which helps to take advantage of the feature extraction capabilities of each model. In other words, the synthesized model can extract features at a deeper level, significantly improving the prediction ability compared to traditional ensemble models.

@@

\verb|NHITS| is designed to target the prediction of long-horizon time-series data. According to \cite{challu2023nhits}, the structure of \verb|NHITS| consists of multiple consecutive stacks. Each stack consists of multiple consecutive blocks. At each block, historical data is used to predict future data and past data. The residual of the previous block is used as input data for the following block. Specifically, at block $l$, with $L$ past data samples ($\mathbf{y}_{t-L:t, l-1}$), the features will be extracted as follows (from \cite{challu2023nhits}):

@@

Accordingly, \verb|Pooling| compresses the original data into $\mathbf{y}_{t-L:t, l}^{(p)}$. \verb|FullyConnected| contains many perceptron layers with nonlinear activation functions. $\mathbf{\theta}_l^f, \mathbf{\theta}_l^b$ are the forecast and backcast interpolation coefficients, which are used to aggregate the output values of block $l$ using the interpolation function $g(\cdot)$. The dimensionality of the interpolation coefficients in each stack is determined by the expressiveness ratio: $\left| \theta_l \right| = \left \lceil r_l H \right \rceil$. Typically, the expressiveness ratio will be very small in the first stacks and increase towards the end. Accordingly, the stacks can simulate frequencies from low to high. In addition, by using $g(\cdot)$, \verb|NHITS| does not require too much hardware to train the neural network in the case of large $H$.

@@

Our method consists of two main parts that work in parallel: (1) - Feature extraction; (2) - Parameter synthesis. The overview of the method is illustrated in \ref{fig:flow}. In the feature extraction part, we combine two types of features from \verb|CNN| and \verb|LSTM| networks. In the parameter synthesis part, we use \verb|MAML| to synthesize the parameters of the models. Due to the contribution of \verb|LSTM| and \verb|CNN| features, we expect to effectively extract hidden features from aperiodic data. By using \verb|MAML| in the weight synthesis process, the proposed method is expected to be a reasonable and effective alternative to traditional ensemble models in minimizing the impact of variance variation, effectively synthesizing external factors, and preserving hidden long-term dependencies in the past.

@@

A data sample consists of pairs of values $(\mathbf{x}_{t-L:t}, y)$. In which, $\mathbf{x}_{t-L:t}$ includes $L$ historical values from time $t$ back; $y\in \{0,1\}$ is the data label, showing the decreasing or increasing trend of the data sample $x_{t+1}$ compared to $x_{t}$. Depending on each problem and the implementation, the elements in $\mathbf{x}_{t-L:t}$ can be vectors or scalar numbers. For example, for stock data, $\mathbf{x}_{t-L:t}$ can contain $L$ data vectors $\vec x_i = (\text{open, low, high, close})$ or just a single close price value.

@@

\begin{align}
    \mathbf{x'}_{t-L:t} &= \mathbf{FullyConnected}\left( \mathbf{x'}_{t-L:t} \right)\\
    \mathbf{h}_{LSTM} &= \mathbf{BidirectionalLSTM}\left( \mathbf{x'}_{t-L:t} \right)\\
    \mathbf{h}_{CNN} &= \mathbf{Convolution1D}\left( \mathbf{x'}_{t-L:t} \right)
    \label{eq:feature}
\end{align}

@@

The \verb|LSTM| network maintains cell-state values to selectively store long-term dependencies. This is very suitable for solving time-series data problems. On the other hand, future values often depend heavily on recent historical values. We propose to use the \verb|CNN| network to emphasize local features, thereby directing part of the model's attention to certain time points. Therefore, the proposed method can not only remember long-term features but also highlight short-term features.

@@

We use \verb|MAML| to train and aggregate the weights of the models at the tasks. As mentioned in \ref{sec.relatedWork}, parameter optimization in the ML approach is to solve the two equations \ref{eq:inner_opt} and \ref{eq:outer_opt} using optimization methods on the support and query data. Specifically, the optimization process includes many global steps (outer optimization), performed on all tasks participating in training. Each global step includes many local steps (inner optimization) performed on each individual task. At global step $r$-th, the $e$-th local optimization process at the support set of task $t$ occurs as follows:

\begin{align}
    \begin{cases}
        \theta_t^{(0)} &= \phi_{r-1} \\
        \theta_t^{(e)} &= \theta_t^{(e-1)} - \alpha \nabla_{\theta} \mathcal{L}^{task}_t\left( \theta_t^{(e-1)}, \mathcal{D}_t^{support} \right)
    \end{cases}
\end{align} In which, $\phi_{r-1}$ is the result of the $r-1$ global optimization process, $\alpha$ is the inner learning rate.

@@

Next, the outer optimization process at the global step is performed by aggregating the losses on the query set of the tasks and optimizing on it (equation \ref{eq:outer_sol}).

@@

The presence of the product of second order derivatives in the equation \ref{eq:outer_sol_} makes the derivation process complicated because it requires a lot of overhead to maintain the Hessian matrices. Therefore, the number of computational steps to find $\theta^*$ needs to be limited. In practice, methods using ML \citep{fallah2020personalized, chen2018federated, nguyen2022meta,finn2017model, li2017meta} often choose $E\in [1,5]$.

@@

Our goal is to use \verb|multi-fx| dataset to extract and aggregate information about outliers (i.e. market, economic, political, etc.) that are believed to influence the outcome of a given financial index \citep{overreactioncontrarian,fama1970efficient,mech1993portfolio}. Meanwhile \verb|USD/JPY| dataset is used to test our hypothesis that future data implicitly depend on certain points in the past and that efficient past feature aggregation is needed to reveal these dependencies.

@@

We compare the proposed method with \verb|NHITS| using the above metrics and dataset. For each dataset, each attribute will be predicted for future trends based on past data. The final result is calculated by averaging the metrics across the attributes. We also test different feature extractors (\verb|LSTM|, \verb|CNN|) for the proposed method to demonstrate the deep data understanding capabilities of \verb|LSTM+CNN|. The overall data is structured into training sets, validation sets, and testing sets with a ratio of 6:2:2 for training, fine-tuning and testing model, respectively.

@@

For \verb|NHITS| model, we split the data as usual according to the above predetermined ratio. For the proposed algorithm, because the meta-training and meta-testing processes require dividing the data into small tasks, and allowing the model to adapt on the support set of each task, we split the data into 40-60 tasks (see table \ref{tab:stat_} for statistical detail). In each task, the support set accounts for 20\% with the purpose of letting the model adapt to the data, the query set accounts for 80\% to check the model's compatibility. We use 50\% of tasks for meta-training, 25\% for meta-validation, and 25\% for meta-testing. With this division, we ensure the fairness that the ML model is trained with the same amount of data as \verb|NHITS|. After the data configuration, we perform experiment as in appendix \ref{app:our_experiment} and \ref{app:nhits_experiment}.

@@

Table \ref{tab:rs} compares the results between the proposed model and the \verb|NHITS| model on the above datasets. On aperiodic data, our method achieves a much higher convergence rate than the \verb|NHITS| model on all metrics. Indeed, the accuracy improves from 5\% to 11\% compared to \verb|NHITS| on \verb|USD/JPY| and \verb|multi-fx| data. The remaining metrics also increase from 5\% to 20\%. On periodic data (\verb|ETT-m2| and \verb|WTH| dataset), our method obtains results that are higher than or close to those of \verb|NHITS|. Specifically, our results differ by 1-3\% on all metrics compared to the results of \verb|NHITS|. In the following subsections, we analyze these results on aperiodic datasets based on two factors: (1) - Feature extraction ability; (2) - Model synthesis ability; and relate to periodic datasets.

@@

We trained three models \verb|CNN|, \verb|LSTM|, \verb|LSTM+CNN| on 2600 samples of \verb|USD/JPY| data in the conventional way. After observing their training process (figure \ref{fig:lstm_cnn}), we found that \verb|CNN| has excellent local feature extraction ability on short-term data. On the other hand, \verb|LSTM| does not improve after 40 epochs, indicating that focusing on exploiting short/long-term dependencies without extracting deep enough features will not yield good results. When combining both features, the model shows an improvement in learning ability on the training set. The convergence rate is generally higher than when using each model separately. This can be explained by the fact that \verb|LSTM| successfully integrated long-term features into the model, as well as taking advantage of the local features of \verb|CNN|.

@@

However, when observing the results on the validation set, all three models above give very poor results. To explain this, we think that although they can learn well on the training set, the models lose their generalization ability because this is a complex type of data, different from data such as images, videos, or periodic data. If the model's adaptability is not high, poor results on the validation set are predictable. However, \textbf{the graph in figure \ref{fig:lstm_cnn} shows that we cannot deny the feature extraction ability of} \verb|LSTM+CNN|.

@@

The feature extraction process of \verb|NHITS| shows that this method is trying to simulate the frequency resolution process of Fourier transform. The frequency features are the most important information for periodic data. Therefore, the predictions of \verb|NHITS| can easily achieve high accuracy on this type of data. However, it is very difficult to resolve the frequency on non-periodic data. Moreover, it is not difficult to see that the feature extraction techniques of \verb|NHITS| are extremely simple (only including \verb|FullyConnected| and \verb|Pooling| classes). These are the reasons why \verb|NHITS| encountered great difficulties on \verb|multi-fx|, \verb|USD/JPY| dataset and non-periodic data in general. On the other hand, our proposed method uses deep learning networks for feature extraction, so it can easily obtain deep hidden features. Therefore, on both periodic and aperiodic data, our method achieves results approximately equal to or better than \verb|NHITS|.

@@

When using \verb|MAML| to synthesize the features of the \verb|LSTM| and \verb|LSTM+CNN| models, the compatibility (indicated by early convergence) as well as the results (indicated by accuracy) of the model using \verb|LSTM+CNN| are significantly higher than those of the model using only \verb|LSTM| (see figure \ref{fig:lstm_cnn_feature}) and \verb|NHITS| model. Indeed, on \verb|multi-fx| dataset, after only 40 epochs, the \verb|LSTM+CNN| features converge to over 60\% and after 100 epochs, the model converges to 62\%, exceeding what the \verb|LSTM| features can do. The variation of the accuracy values is very small, indicating the stability of \verb|MAML|. For the ML model using only \verb|CNN|, the accuracy and error values do not even improve during the entire training process. We explain this by saying that the features of \verb|CNN| focus too much on the locality of the data and ignore the long-term dependencies, which is an important feature of time-series data. In contrast, the features of \verb|LSTM+CNN| not only capture the long-term dependencies but also highlight the local properties in the data, making it completely outperform \verb|CNN| as well as \verb|NHITS|. Therefore, \textbf{we conclude that ML is more flexible and efficient in synthesizing sub-models than traditional rigid ensemble models}.

@@

\appendix
\section{Experimental detail\\for proposed method}
\label{app:our_experiment}

>>>>

    \textbf{Abstract}. Predicting aperiodic time-series data (e.g. stock price, foreign exchange, Bitcoin price, etc.) is a difficult task for machine learning models because this type of data has high and not stationary variance; does not show clearly periodicity, making it difficult to extract features; depends not only on past values but also on external factors such as economic and political situations, which make them unstable and non-cyclical. To overcome the above challenges, we use Meta-learning to train a combined \verb|LSTM| and \verb|CNN| network, thereby effectively extracting and synthesizing deep hidden features of the data over time. Experiments on aperiodic data (foreign exchange data of 60 currency pairs over 24 years from 2000 to 2024) show that the proposed method performs well and has 5\%-11\% higher accuracy compared to the \verb|NHITS| - the state-of-the-art model in 2023 on time-series data, in the task of predicting the trend (upward or downward) of the next trading day. Meanwhile, the results obtained on periodic data (Electricity Transformer Temperature, Weather) are approximately \verb|NHITS|.

@@

Ensemble models \citep{sadeghi2021combined, zafeiriou2020intraday, ali2020complete} are often used to mitigate the effects of variance variation challenge. Ensemble models provide a holistic and multi-perspective view based on sub-models, thereby helping the overall model adapt to strong variance changes. However, under the rigid synthesis of traditional ensemble models, the results obtained on non-periodic datasets are very poor.

@@

To overcome the pattern variation challenge, most studies use features extracted from Long Short-Term Memory neural network (\verb|LSTM|) \cite{hochreiter1997long}, Artificial Neural Network (\verb|ANN|), and Convolution Neural Network (\verb|CNN|) \cite{lecun1989handwritten}. Specifically, in 2022, 20\% of all publications related to financial index prediction used \verb|LSTM|, 20\% used \verb|ANN|, and 6\% used \verb|CNN| \cite{ayitey2023forex}. This is completely understandable because \verb|LSTM| and \verb|CNN| have proven their ability to extract spatial and temporal features effectively. However, when faced with aperiodic data with only 2600 samples, after 40 training steps, our experiment shows that \verb|LSTM| is still underfitting, while \verb|CNN| is overfitting after only 5 training steps.

@@

Regarding the challenge of external influences, \cite{fama1970efficient} hypothesizes that different time-series datasets of the same domain at the same time reflect the impact of external factors. Studies \citep{overreactioncontrarian, mech1993portfolio} also show the dependence between the financial ratios of a given company and the ones of other companies. This further strengthens the hypothesis in study \cite{fama1970efficient}. Therefore, by exploiting data sources with the same domain, we can capture information reflecting external factors.

@@

We approach the above challenges through the combination of feature and model synthesis. Specifically, by using Meta-learning (ML) \cite{finn2017model}, we efficiently synthesize the parameters of local models, which significantly reduces variance loss and increases the ability to collect information reflecting external factors from datasets in the same domain. To effectively learn laten patterns, we propose a method that combines features from \verb|CNN| and \verb|LSTM|. In addition, we assume that aperiodic time-series data also has hidden long-term dependencies. Traditional approaches use a fixed amount of past data (lookback window) to train the model. This poses a great obstacle to the learning process because long-term features will be forgotten over time. On the other hand, ML divides the dataset into many parts to learn and synthesize the learned parameters effectively, so it can handle this challenge well.

@@

Finally, we demonstrate the superiority of the proposed algorithm by solving the problem of predicting the trend (upward or downward) of time-series data and comparing the results with the existing state-of-the-art (SOTA) model (\verb|NHITS| \cite{challu2023nhits}) on two types of data: (1) - Aperiodic data, which contains USD/JPY exchange rate data and the exchange rate data of 60 currency pairs, comprising 18 countries from 2000 to 2024; (2) - Periodic data which contains Electricity Transformer Temperature \cite{zhou2021informer} and Weather \cite{Kolle} dataset. These datasets are publicly available on the Internet and can be downloaded easily. The official implementation can be found at \url{https://github.com/baolongnguyenmac/multi_fx}.

@@

In summary, our main contributions include the following: First, we present a novel \textbf{feature combination technique} for aperiodic time-series data that extracts and combines features using both \verb|LSTM| and \verb|CNN| models. Second, we discover the \textbf{hidden long-term dependencies} and demonstrate experimentally that aperiodic time-series data not only rely on external factors but also exhibit dependencies with past period. Third, we introduce an \textbf{efficient model parameter aggregation approach} by employing ML algorithm in place of traditional ensemble models for combining parameters from multiple machine learning models. Finally, we validate our method through experiments on exchange rate datasets and show its effectiveness in comparison to a SOTA model, \verb|NHITS|.

@@

As aforementioned, \verb|LSTM| is a well-known neural network for handling prediction problems on time-series data. \verb|LSTM| is commonly used because it efficiently handles the problem of vanishing gradients (easily encountered when using Recurrent neural networks) and can effectively exploit nonlinear relationships in data. Indeed, by maintaining the cell-state in each iteration, \verb|LSTM| can overcome the vanishing gradient problem, thereby preserving the ability to capture long-term dependencies \cite{cheng2018applied}. In addition, \verb|LSTM| performs feature extraction with nonlinear activation functions, which helps the model parameters capture the nonlinearity of the data \cite{he2016exploiting}. These factors make \verb|LSTM| to be the first choice to think of when solving problems on time-series data.

@@

Hybrid ensemble models have been widely used in time-series processing problems and have been experimentally proven to be more accurate than standard time-series models because they can synthesize the strengths of many sub-models \cite{ayitey2023forex}. However, current ensemble model synthesis forms are still very rigid because they can only synthesize based on the final results (e.g. voting mechanism of bagging models) and near-final results (e.g. stacking models). From the perspective of ensemble models, the equation \ref{eq:outer_opt} can be considered an optimization-based method of synthesizing sub-models, which helps to take advantage of the feature extraction capabilities of each model. In other words, the synthesized model can extract features at a deeper level, significantly improving the prediction ability compared to traditional ensemble models.

@@

\verb|NHITS| is designed to target the prediction of long-horizon time-series data. According to \cite{challu2023nhits}, the structure of \verb|NHITS| consists of multiple consecutive stacks. Each stack consists of multiple consecutive blocks. At each block, historical data is used to predict future data and approximate past data. The difference between historical data and the approximate one of the previous block is used as input data for the following block. Specifically, at block $l$, with $L$ past data samples ($\mathbf{y}_{t-L:t, l-1}$), the features will be extracted as follows (from \cite{challu2023nhits}):

@@

Accordingly, \verb|Pooling| compresses the original data into $\mathbf{y}_{t-L:t, l}^{(p)}$. \verb|FullyConnected| contains many perceptron layers with nonlinear activation functions. $\mathbf{\theta}_l^f, \mathbf{\theta}_l^b$ are the forecast and backcast interpolation coefficients, which are used to aggregate the output values of block $l$ using the interpolation function $g(\cdot)$. The dimensionality of the interpolation coefficients in each stack is determined by the expressiveness ratio: $\left| \theta_l^f \right| = \left \lceil r_l H \right \rceil$. Typically, the expressiveness ratio will be very small in the first stacks and increase towards the end. Accordingly, the stacks can simulate frequencies from low to high. In addition, by using $g(\cdot)$, \verb|NHITS| does not require too much hardware to train the neural network in the case of large $H$.

@@

Our method consists of two main parts that work in parallel: (1) - Feature extraction; (2) - Parameter synthesis. The overview of the method is a ML process as mentioned in section \ref{sec.relatedWork} and is illustrated in figure \ref{fig:flow}. In the feature extraction part, we combine two types of features from \verb|CNN| and \verb|LSTM| networks. In the parameter synthesis part, we use \verb|MAML| to synthesize the parameters of the models. Due to the contribution of \verb|LSTM| and \verb|CNN| features, we expect to effectively extract hidden features from aperiodic data. By using \verb|MAML| in the weight synthesis process, the proposed method is expected to be a reasonable and effective alternative to traditional ensemble models in minimizing the impact of variance variation, effectively synthesizing external factors, and preserving hidden long-term dependencies in the past.

@@

A data sample consists of pairs of values $(\mathbf{x}_{t-L:t}, y)$. In which, $\mathbf{x}_{t-L:t}$ includes $L$ historical values from time $t$ back; $y\in \{0,1\}$ is the data label, showing the decreasing or increasing trend of the data sample $x_{t+1}$ compared to $x_{t}$. Depending on each problem and the implementation, the elements in $\mathbf{x}_{t-L:t}$ can be a matrix or a vector. For example, for stock data, $\mathbf{x}_{t-L:t}$ can contain $L$ vectors $\vec x_i = (\text{open, low, high, close})$ or can be a vector of close price values only.

@@

\begin{align}
    \mathbf{x'}_{t-L:t} &= \mathbf{FullyConnected}\left( \mathbf{x}_{t-L:t} \right)\\
    \mathbf{h}_{LSTM} &= \mathbf{BidirectionalLSTM}\left( \mathbf{x'}_{t-L:t} \right)\\
    \mathbf{h}_{CNN} &= \mathbf{Convolution1D}\left( \mathbf{x'}_{t-L:t} \right)
    \label{eq:feature}
\end{align}

@@

The \verb|LSTM| network maintains cell-state values to selectively store long-term dependencies. This is very suitable for solving time-series data problems. On the other hand, future values often depend heavily on recent historical values. We propose to use the \verb|CNN| network to emphasize local features, thereby directing part of the model's attention to certain time points. Therefore, the proposed method not only remembers long-term features but also highlights short-term features.

@@

We use \verb|MAML| to train and aggregate the weights of the models at the tasks. As mentioned in \ref{sec.relatedWork}, parameter optimization in the ML approach is to solve the two equations \ref{eq:inner_opt} and \ref{eq:outer_opt} using optimization methods on the support and query data. Specifically, the optimization process includes many global steps (outer optimization), performed on all tasks participating in training. Each global step includes many local steps (inner optimization) performed on each individual task. At outer optimization step $r$-th, the $e$-th inner optimization step at the support set of task $t$ occurs as follows:

\begin{align}
    \begin{cases}
        \theta_t^{(0)} &= \phi_{r-1} \\
        \theta_t^{(e)} &= \theta_t^{(e-1)} - \alpha \nabla_{\theta} \mathcal{L}^{task}_t\left( \theta_t^{(e-1)}, \mathcal{D}_t^{support} \right)
    \end{cases}
\end{align} In which, $\phi_{r-1}$ is the result of the $r-1$ outer optimization process, $\alpha$ is the inner learning rate.

@@

Next, the outer optimization process is performed by aggregating the losses on the query set of the tasks and optimizing on it (equation \ref{eq:outer_sol}).

@@

The presence of the product of second order derivatives in the equation \ref{eq:outer_sol_} makes the derivation process complicated because it requires a lot of overhead to maintain Hessian matrices. Therefore, the number of computational steps to find $\theta^*$ needs to be limited. In practice, methods using ML \citep{fallah2020personalized, chen2018federated, nguyen2022meta,finn2017model, li2017meta} often choose $E\in [1,5]$.

@@

Our goal is to use \verb|multi-fx| dataset to check the information extraction and aggregation of external factors (i.e. market, economic, political, etc.) that are believed to influence the outcome of a given financial index \citep{overreactioncontrarian,fama1970efficient,mech1993portfolio}. Meanwhile \verb|USD/JPY| dataset is used to test our hypothesis that future data implicitly depend on certain points in the past and that efficient past feature aggregation is needed to reveal these dependencies.

@@

We compare the proposed method with \verb|NHITS| using the above metrics and dataset. For each dataset, each attribute will be predicted the future trends based on past data. The final result is calculated by averaging the metrics across the attributes. We also test different feature extractors (\verb|LSTM|, \verb|CNN|) for the proposed method to demonstrate the deep data understanding capabilities of \verb|LSTM+CNN|.

@@

For \verb|NHITS| model, data is structured into training sets, validation sets, and testing sets with a ratio of 6:2:2 for training, fine-tuning and testing model, respectively. For the proposed algorithm, because the meta-training and meta-testing processes require dividing the data into small tasks, and allowing the model to adapt on the support set of each task, we split the data into 40-60 tasks (see table \ref{tab:stat_} for statistical detail). In each task, the support set accounts for 20\% with the purpose of letting the model adapt to the data, the query set accounts for 80\% to check the model's compatibility. We use 50\% of tasks for meta-training, 25\% for meta-validation, and 25\% for meta-testing. With this division, we ensure the fairness that the ML model is trained with the same amount of data as \verb|NHITS| (see appendix \ref{app:data}). The detail of experiments can be found at appendix \ref{app:our_experiment} and \ref{app:nhits_experiment}.

@@

Table \ref{tab:rs} compares the results between the proposed model and the \verb|NHITS| model using the above datasets. On aperiodic data, our method achieves a much higher convergence rate than the \verb|NHITS| model on all metrics. Indeed, the accuracy improves from 5\% to 11\% compared to \verb|NHITS| on \verb|USD/JPY| and \verb|multi-fx| data. The remaining metrics also increase from 5\% to 20\%. On periodic data (\verb|ETT-m2| and \verb|WTH| dataset), our method obtains results that are higher than or close to those of \verb|NHITS|. Specifically, our results differ by 1-3\% on all metrics compared to the results of \verb|NHITS|. In the following subsections, we analyze these results on aperiodic datasets based on two factors: (1) - Feature extraction ability; (2) - Model synthesis ability; and relate to periodic datasets.

@@

We trained three models \verb|CNN|, \verb|LSTM|, \verb|LSTM+CNN| on 2600 samples of \verb|USD/JPY| data in the conventional way. After observing their training process (figure \ref{fig:lstm_cnn}), we found that \verb|CNN| has excellent local feature extraction ability on short-term data. On the other hand, \verb|LSTM| does not improve after 40 epochs, indicating that focusing on exploiting short/long-term dependencies without extracting deep enough features will not yield good results. When combining both features, the model shows an improvement in learning ability on the training set. The convergence rate is generally higher than when using each model separately. This can be explained by the fact that \verb|LSTM| \textbf{successfully integrated long-term features into the model, as well as taking advantage of the local features highlighting ability of} \verb|CNN|.

@@

However, when observing the results on the validation set, all three models above give very poor results. To explain this, we think that although they can learn well on the training set, the models lose their generalization ability because this is a complex type of data, which is different from data such as images, videos, or periodic data. If the model's adaptability is not good, poor results on the validation set are predictable. This suggests that there is a need for a method that efficiently utilizes and synthesizes the obtained features, which is discuss in detail in the next subsection.

@@

The feature extraction process of \verb|NHITS| shows that this method is trying to simulate the frequency resolution process of Fourier transform. The frequency features are the most important information for periodic data. Therefore, the predictions of \verb|NHITS| can easily achieve high accuracy on this type of data. However, it is very difficult to resolve the frequency on non-periodic data. Moreover, it is not difficult to see that the feature extraction techniques of \verb|NHITS| are extremely simple (only including \verb|FullyConnected| and \verb|Pooling| layers). These are the reasons why \verb|NHITS| encountered great difficulties on \verb|multi-fx|, \verb|USD/JPY| dataset and non-periodic data in general. On the other hand, our proposed method uses deep learning networks for feature extraction, so it can easily obtain deep hidden features. Therefore, on both periodic and aperiodic data, our method achieves results close to or better than \verb|NHITS|.

@@

When using \verb|MAML| to synthesize the features of the \verb|LSTM| and \verb|LSTM+CNN| models, the compatibility (indicated by early convergence) as well as the results (indicated by accuracy) of the model using \verb|LSTM+CNN| are significantly higher than those of the model using only \verb|LSTM| (see figure \ref{fig:lstm_cnn_feature}) and \verb|NHITS| model. Indeed, on \verb|multi-fx| dataset, after only 40 epochs, the \verb|LSTM+CNN| features converge to over 60\% and after 100 epochs, the model converges to 62\%, exceeding what the \verb|LSTM| features can do. The variation of the accuracy values is very small, indicating the stability of \verb|MAML|. For the ML model using only \verb|CNN|, the accuracy and error values do not even improve during the entire training process. To explain this, we argue that the features of \verb|CNN| have focused too much on the locality of the data and ignored the long-term dependencies, which is an important feature of time-series data. In contrast, the features of \verb|LSTM+CNN| not only capture the long-term dependencies but also highlight the local properties in the data, making it completely outperform \verb|CNN| as well as \verb|NHITS|. Therefore, \textbf{we conclude that ML is more flexible and efficient in synthesizing sub-models as well as external factors than traditional rigid ensemble models}.

@@

\appendix
\section{Fairness in data}
\label{app:data}

% Với \verb|NHITS|, chúng tôi sử dụng các chia dữ liệu truyền thống. Tập dữ liệu ban đầu được chia thành training set, validation set, và testing set theo tỷ lệ 6:2:2.

For \verb|NHITS|, we use the traditional data split. The dataset is split into a training set, validation set, and testing set in a 6:2:2 ratio.

% Đối với phương pháp đề xuất, chúng tôi chia dữ liệu thành các tập dữ liệu nhỏ, mỗi tập dữ liệu lại được chia thành tập support và tập query. Gọi số lượng task sau khi chia là $n$, chúng tôi sử dụng $0.5n$ tasks cho meta-training, $0.25n$ tasks cho meta-validating, và $0.25n$ tasks cho meta-testing. Trong mỗi task, tập support chiếm 20\%, tập query chiếm 80\%.

For the proposed method, we split the data into small datasets, each of which is divided into a support set and a query set. Let the number of tasks after splitting be $n$, we use $0.5n$ tasks for meta-training, $0.25n$ tasks for meta-validating, and $0.25n$ tasks for meta-testing. In each task, the support set accounts for 20\%, and the query set accounts for 80\%.

% Gọi $M$ là số mẫu dữ liệu trong tập dữ liệu ban đầu, số mẫu dữ liệu sử dụng để train, validate, và test cho \verb|NHITS| lần lượt là $0.6M, 0.2M, 0.2M$. Đối chiếu với lượng dữ liệu sử dụng cho phương pháp đề xuất, mỗi task sẽ bao gồm $\frac{M}{n}$ mẫu dữ liệu. Chúng tôi sử dụng $0.5M$ cho meta-training (vì số task cho meta-training là $0.5n$). Tiếp đến, trong quá trình validate hoặc test, chúng tôi sử dụng 20\% dữ liệu mỗi task cho quá trình model adaptation và 80\% dữ liệu còn lại cho validate hoặc test. Theo đó, tổng số mẫu dùng cho quá trình huấn luyện và adaptation là $0.5M + 0.25n \times 0.2\frac{M}{n} + 0.25n \times 0.2\frac{M}{n} = 0.6M$. Số mẫu dùng để đánh giá mô hình trong quá trình validation và testing đều là $0.25n \times 0.8\frac{M}{n} = 0.2M$. Như vậy, chúng tôi đã bảo đảm sự công bằng trong huấn luyện và kiểm thử cho các thí nghiệm.

Let $M$ be the number of data samples in the original dataset, the number of data samples used for training, validation, and testing for \verb|NHITS| are $0.6M, 0.2M, 0.2M$ respectively. Compared to the amount of data used for the proposed method, each task will include $\frac{M}{n}$ data samples. We use $0.5M$ samples for meta-training (because the number of tasks for meta-training is $0.5n$). Next, during the validation or testing process, we use 20\% of the data of each task for the model adaptation process and the remaining 80\% of the data for validation or testing. Accordingly, the total number of samples used for training and adaptation is $0.5M + 0.25n \times 0.2\frac{M}{n} + 0.25n \times 0.2\frac{M}{n} = 0.6M$. The number of samples used to evaluate the model in the validation and testing process is both $0.25n \times 0.8\frac{M}{n} = 0.2M$. Thus, we have ensured fairness in training and testing for the experiments.

\section{Experimental detail\\for proposed method}
\label{app:our_experiment}
